---
title: "Getting and Cleaning Data Course Project"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Merges the training and the test sets to create one data set.
## 1.1 Download and Unzip files
```{r}
# Download zip file
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl, destfile = "./data/DataSet.zip", method = "curl")

# Unzip files
unzip(zipfile = "./data/DataSet.zip", exdir = "./data")
```


## 1.2 Read Files
```{r}
# Read training files
x_train <- read.table("./data/UCI HAR Dataset/train/X_train.txt")
y_train <- read.table("./data/UCI HAR Dataset/train/y_train.txt")
subject_train <- read.table("./data/UCI HAR Dataset/train/subject_train.txt")

# Read test files
x_test <- read.table("./data/UCI HAR Dataset/test/X_test.txt")
y_test <- read.table("./data/UCI HAR Dataset/test/y_test.txt")
subject_test <- read.table("./data/UCI HAR Dataset/test/subject_test.txt")

# Read features file
features <- read.table("./data/UCI HAR Dataset/features.txt", 
                       stringsAsFactors = FALSE)

# Read Activity name file
activityName <- read.table("./data/UCI HAR Dataset/activity_labels.txt")
colnames(activityName) <- c("activityID", "activityType")
```


## 1.3 Merge train and test data
```{r}
trainData <- cbind(subject_train, y_train, x_train)
testData <- cbind(subject_test, y_test, x_test)
fullData <-rbind(trainData, testData)
```


## 2. Extracts only the measurements on the mean and standard deviation for each measurement
```{r}
featureIndex <- grep(("mean|std"), features[,2])
MeanStdData <- fullData[ ,c(1, 2, featureIndex + 2)]
colnames(MeanStdData) <- c("subject", "activityID", features[featureIndex,2])
```


## 3. Uses descriptive activity names to name the activities in the data set
```{r}
finalData <- merge(MeanStdData, activityName, by = "activityID", all.x = T)
```

## 4. Appropriately labels the data set with descriptive variable names
```{r}
names(finalData) <- gsub("\\()","", names(finalData))
names(finalData) <- gsub("^t", "Time", names(finalData))
names(finalData) <- gsub("^f", "Frequency", names(finalData))
names(finalData) <- gsub("-mean", "Mean", names(finalData))
names(finalData) <- gsub("-std", "Std", names(finalData))
names(finalData) <- gsub("Acc", "Accelerometer", names(finalData))
names(finalData) <- gsub("Gyro", "Gyroscope", names(finalData))
```


## 5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r}
library(dplyr)
groupData <- finalData %>%
    group_by(subject, activityID) %>%
    summarise_each(funs(mean))
write.table(groupData, "./data/TidyMeanData.txt", row.names = T)
```










